---
title: "Coverage analysis of indentified species"
author: "Charles"
output:
  html_document:
    toc: true 
    toc_depth: 3
    number_sections: true  
    theme: united 
    highlight: tango
    df_print: paged
---

In this notebook, we will produce genome coverage plots for the identified species

# Load libraries
```{r, include=FALSE, cache=FALSE}
#.libPaths("/home/groups/astraigh/R_packages_3.6")
library(tidyr)
library(dplyr)
library(ggplot2)
```


# Covaerage analysis

## Quick look
Load the coverage report in a tibble
```{r}
bp_coverage <- read.table('/scratch/groups/astraigh/genomics_minicourse/shared/data/mock12/alignments/ont.coverage.txt',sep='\t', col.names=c("scaffold","coverage","N_bp","scaffold_size","percent_bp"))
bp_coverage <-as_tibble(bp_coverage)
```

Take a look
```{r}
bp_coverage
```

Plot the coverage distribution (percent of basepairs as a function of the coverage depth)
```{r}
bp_coverage %>%
  dplyr::filter(coverage>0) %>%
  ggplot(aes(x=coverage, y=percent_bp, color=scaffold))+
  geom_point(size=1)+
  theme_minimal()+
  scale_x_log10()+
  scale_y_log10()+
  annotation_logticks()
```


## Clean up
Some species have multiple scaffolds. We need to group together scaffolds from the same species. 

Firs, we load the scaffold to taxid map.
```{r}
scaf2taxid <- as_tibble(read.table('/scratch/groups/astraigh/kraken/standard/scaffold2taxid.map',sep=' ', col.names=c("scaffold","taxid")))
head(scaf2taxid)
```

We add species info
```{r}
genomes_info <- as_tibble(read.table('/scratch/groups/astraigh/genomics_minicourse/shared/data/mock12/kraken/top19.genomes.txt',sep='\t', col.names=c("taxid","species_taxid","name","file","blank","species"))) %>%
   dplyr::select(-file, -blank)

scaffold_withInfo<- scaf2taxid %>%
  inner_join(genomes_info, by='taxid')

scaffold_withInfo
```

Now we can group together the coverage values by species
```{r}
bp_coverage_byspecies <-
  bp_coverage %>% 
  left_join(scaffold_withInfo %>% dplyr::select(scaffold, species)) %>%
  group_by(species, coverage) %>%
  summarise(N_bp=sum(N_bp), scaffold_size=sum(scaffold_size)) %>%
  group_by(species) %>%
  mutate(percent_bp=N_bp/scaffold_size) %>%
    ungroup()
```

Take a look at the result
```{r}
bp_coverage_byspecies
```


### Coverage distribution by species

```{r, fig.width=12, fig.height=4}
bp_coverage_byspecies %>%
  dplyr::filter(coverage>0) %>%
  ggplot(aes(x=coverage, y=percent_bp, color=species))+
  geom_point(size=1)+
  theme_minimal()+
  scale_x_log10()+
  scale_y_log10()+
  annotation_logticks()
```


### Mean coverage
Get the mean coverage by species
```{r}
bp_coverage_byspecies_mean <-
  bp_coverage_byspecies %>%
  group_by(species) %>%
  summarise(coverage_mean=sum(coverage*(N_bp/sum(N_bp)))) %>%
  ungroup()

bp_coverage_byspecies_mean
```

And lastly, plot.
```{r}
bp_coverage_byspecies_mean %>%
  arrange(desc(coverage_mean)) %>%
  mutate(species=as.character(species)) %>%
  mutate(species=factor(species, levels=species)) %>%
  ggplot(aes(x=species, y=coverage_mean)) +
  geom_bar(stat="identity")+
  theme_minimal()+
  theme(axis.text.x = element_text(angle = 90))
```

